%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{figr-showcase}

figr showcase
---------

`figr` is designed to be a simple figure and table reference manager. This vignette shows how to use the package. 

First, create a figure:
```{r message=FALSE, warning=FALSE}
require(ggplot2)
p = qplot(rnorm(500),geom="histogram")
```

Now we use figr to index it:
```{r message=FALSE}
require(figr)
addFigure('example', p, caption='a histogram')
# note that figr makes a copy of the plot object
rm(p)
```
Cross-referencing can be achieved using the HTML anchor tag (`<a>`). We can define a new hook for anchoring a code block.

```{r results='hide', message=FALSE}
require(knitr)
knit_hooks$set(anchor = function(before, options, envir) {
    if (before) 
	  paste('<a name="', options$anchor, '"></a>', sep='')
})
```

Note that the anchors will be placed *above* the code chunk, because we are using `if(before)` to specify that the hook is evaluated before the code chunk. (We can place the anchor below the chunk by using `if(!before)`). We can automatically generate an anchor that will match the figure citation with `anchorFigure('example')` (note that we can create the anchor before creating the figure). We will now place the figure. Note that the code chunk below includes the option `anchor=anchorFigure('example')`.

```{r message=FALSE, warning=FALSE, anchor=anchorFigure('example')}
placeFigure('example')
```

`figr` determines figure numbers by ranking them based on the order they are placed or cited in the document. We can reference `r citeFigure("example")` (I just did!) by typing `citeFigure("example")`.  

Calling `placeFigure('example')` a second time will not change the rankings. This is the result of code that allows a figure to be cited before it is placed. A figure is not ranked until it is placed or cited, so if for some reason you need to refer to a later figure you must cite or place any intervening figures first. 

```{r message=FALSE, warning=FALSE}
# note you don't actually have to create a figure to cite it!
invisible(citeFigure('ex2'))
```

We'll get to `r citeFigure('ex3')` in a second, but first look at `r citeFigure('ex2')`:

```{r message=FALSE, warning=FALSE, anchor=anchorFigure('ex2')}
p2 = qplot(rlnorm(500),geom="histogram")
addFigure('ex2', p2, 'Another plot.')
placeFigure('ex2')
```

Okay, now for `r citeFigure('ex3')`:

```{r message=FALSE, warning=FALSE, anchor=anchorFigure('ex3')}
p3 = qplot(rgamma(500, 3), geom='histogram')
addFigure('ex3', p3, 'A third plot.')
placeFigure('ex3')
```

Note that automatic generation of a list of figures is not possible (at least, not at the start of a document--but we could create a function to place one at the end of the document). If you are creating a document that requires those features, you may want to consider using LaTeX instead of Markdown.

We can also add a hook for figure captions:

```{r results='hide', message=FALSE}
require(markdown)
knit_hooks$set(plot = function(x, options) {
  paste('<figure><img src="',
        opts_knit$get('base.url'), paste(x, collapse = '.'),
        '"><figcaption>', renderMarkdown(text=options$fig.cap), '</figcaption></figure>',
        sep = '')
})
opts_knit$set(eval.after='fig.cap')
```

A figure caption can be referenced using `labelFigure`. Note the use of the `markdown::renderMarkdown` function inside the plot hook for link formatting. Having the link in the caption is not really useful, but it is nice for labelling consistency.

In `r citeFigure('ex4')` below, the chunk options `fig.cap=labelFigure('ex4')` and `anchor=anchorFigure('ex4')` are used.
```{r anchor=anchorFigure('ex4'), fig.cap=labelFigure('ex4'), message=FALSE}
addFigure('ex4', qplot(rchisq(500, 3), geom='histogram'), 
          'I can define the caption in the code chunk!')
placeFigure('ex4')
```

The package provides identical support for tables (e.g. `addTable`, `citeTable`, etc.) for e.g. kable objects, as shown by `r citeTable('example')`. Note that tables can have identical names to figures without risk of overwriting.

```{r message=FALSE, warning=FALSE, results='asis', anchor=anchorTable('example')}
# note only the data is passed to addTable
d = head(iris[,1:3])
addTable('example', d, caption='a table.')
# code chunk option: results='asis'
kable(placeTable('example'))
```

There is currently no hook for a table caption code chunk option, but the`labelTable` function is defined in anticipation of such functionality.

Things on my TODO list:
* have `figr` check that an anchor has not already been created (what happens when you define duplicate anchors in a page?)
* Figure out table captions (nest a table and caption in a 2-cell table?)
